---
title: "Unveiling the 2022 Iran Protests: Insights from HRANA's Dataset"
subtitle: "Analyzing Civil Unrest, Government Crackdowns, and Public Dissent Through Statistical Estimates"
author: Mehrnoush Mohammadi
thanks: "Code and data are available at: https://github.com/mehrnoush68/WomenLifeFreedom.git"
date: today
date-format: long
abstract: "The 2022 Iran protests were a significant period of civil unrest that brought attention to social inequality, political repression, and human rights violations. However, limited official data on the protests is available, which means we have to rely on estimates from organizations like the Human Rights Activists News Agency (HRANA). In this paper, we have analyzed a dataset compiled from HRANA's reports to provide insights into the dynamics of the protests, the extent of government crackdowns, and the scale of public discontent."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
library(ggplot2)
library(dplyr)
library(arrow)
library(kableExtra)
library(knitr)
library(readr)


#### Read data ####
protest_data <- read_parquet("../data/analysis_data/cleaned_protest_data.parquet")

```


# Introduction

On 16 September 2022, Mahsa Amini, a 22-year-old woman died under police custody in Iran. Reportedly, she was arrested because of not wearing her hijab (headscarf) properly. While officials claimed she died from a heart attack, her family disputed this. CCTV footage showed her collapsing at a "re-education" center. Iranians, especially on social media, demanded justice using hash tags like #MahsaAmini. The incident highlighted concerns about women's rights and government oppression in Iran, drawing international attention and calls for accountability [@CNN2022].

There have been protests in Iran due to various issues such as gender inequality, political oppression, and economic hardships. These issues were further amplified by the tragic death of Mahsa Amini. The protesters, under the banner of "Zan, Zendegi, Azadi" (Women, Life, Freedom), expressed their dissatisfaction with the government. However, there is no official data about the number of casualties, arrests, and other important metrics. To analyze the dynamics of the protests, the government's response, and the extent of public discontent, this paper uses data compiled by the Human Rights Activists News Agency (HRANA) [@justin_oh_2022].

Result paragraph??

To further understand "Women Life, Freedom" movement, in [Data] and [Result], I talk about the nature of the data obtained and analyze the results garnered from the data with suitable tables and charts. Next, [Discussion] provides further insights and future areas of study. Finally, [Conclusion] summarizes my main findings. 

The graphs and tables in this paper were created in R Studio using R [@citeR] and the analysis in a Quarto document. The analysis was conducted with the use of the `ggplot` [@ggplotRef], `tiddyverse` [@tidyverseRef], `knitr` [@knitrRef], `readr` [@readRef], `kableExtra` [@kableextraRef], `janitor` [@janitorRef] and `dyplr` [@dplRef ] packages.

# Data {#sec-data}

The dataset used in this analysis comprises estimates compiled by HRANA [@justin_oh_2022], documenting various aspects of the 2022 Iran protests. This includes statistics on casualties, arrests, protests, and involvement across different cities and universities. While these estimates provide valuable insights, it's important to note that they may not capture the full scope of the protests due to limitations in data collection methods.

```{r}
#| echo: false
#| include: true
#| warning: false

# Generate a sequence of dates within the specified range
start_date <- as.Date("2022-09-16")
end_date <- as.Date("2023-02-21")
date_sequence <- seq(start_date, end_date, by = "day")

# Sample 10 random dates from the sequence
random_dates <- sample(date_sequence, 10)
random_dates <- as.character(random_dates)

# Selecting columns
selected_columns <- protest_data[protest_data$Date %in% random_dates, 
                                 c("Date", "Protesters.Killed", "Protests", "Universities", "Involved.Cities", "Arrested")]

kable(selected_columns, 
      format = "markdown", 
      align = "c",
      caption = "Dataset Compiled By HRANA",
      label = "tbl-protest_data")|>
  kable_styling(latex_options = c("striped", "hold_position"), bootstrap_options = "striped", full_width = FALSE)


```

```{r}
#| label: fig-protests-protesters-killed
#| include: true
#| fig-cap: Number of Protests and Protesters Killed In 158 Days
#| fig-subcap: [ Number of Protests, Number of Protesters Killed ]
#| layout-ncol: 2
#| warning: false
#| message: false
#| echo: false

ggplot(protest_data, aes(x = Date, y = Protests)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Date", y = "Number of Protests")

ggplot(protest_data, aes(x = Date, y = Protesters.Killed)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Date", y = "The total Number of Protestors' Death")

```


# Results

According to the analysis, there has been a continuous increase in the number of protests over a period of 158 days, which signifies that the public's dissatisfaction has been persistent. Moreover, the dataset shows that the intensity of protests has varied significantly across different cities and universities, indicating regional differences within the protest movement.

```{r}
#| include: true
#| label: fig-arrests-detainees
#| fig-cap: "Cumulative Statistics of Arrests and Detainees"
#| warning: false
#| message: false
#| echo: false

ggplot(protest_data, aes(x = as.Date(Date, format="%m/%d/%Y"))) +
  geom_area(aes(y = Arrested, fill = "Individuals Arrested")) +
  geom_area(aes(y = Detainees, fill = "Detainees Identified")) +
  theme_minimal() +
  labs(x = "Date", y = "Cumulative Count") +
  scale_fill_manual(values = c("Individuals Arrested" = "gray", "Detainees Identified" = "darkgray"))

```


# Discussion

## First discussion point {#sec-first-point}

The results emphasize the extent and persistence of public opposition during the 2022 protests in Iran. This is evident from the growing frequency of protests and the cumulative number of arrests and detainees. However, it is necessary to conduct a more in-depth analysis to understand the root causes behind these trends and their possible impact on future socio-political developments in Iran. 

## 2nd discussion point

The incident highlighted longstanding concerns about gender inequality and human rights violations in Iran, particularly regarding women's rights and the government's heavy-handed enforcement of moral codes. Calls for accountability and an end to violence against women resonated both within Iran and on the global stage, underscoring the significance of Amini's death as a catalyst for broader social and political activism.

## Third discussion point

## Weaknesses and next steps

## Limitations and Future Research

Although the HRANA dataset offers valuable insights, it has limitations in its data collection methods. Future research should improve data quality and expand the scope of analysis to include additional socio-economic and political variables.



\newpage


# References


